Compararison of analytical and artificial neural network models in predicting hot deformation behavior of AISI P20 and interpolation-extrapolation capacity validation

### 2022-11-17 Article.tex
Update of the Final part of the paper

\newcommand{\w}{\mbox{\large\ensuremath{\mathsf{w}}}}
\newcommand{\dotp}{\boldsymbol{\cdot}}
%\newcommand{\ccirc}{\kern0.5ex\vcenter{\hbox{$\scriptstyle\circ$}}\kern0.5ex}
%\newcommand{\e}[1]{\exp\left({#1}\right)}
\newcommand{\e}[1]{{\rm e}^{#1}}
\newcommand{\lay}[1]{^{(#1)}}
\newcommand{\mdot}[1]{\accentset{\mbox{\bfseries .}}{#1}}
\newcommand*{\ie}{\emph{i.e.}\@\xspace}
\newcommand*{\versus}{\emph{vs.}\@\xspace}
\newcommand*{\eal}{et \emph{al.}\@\xspace}
\newcommand*{\eg}{e.g.,\@\xspace}
\newcommand{\RMSE}{\text{E}_\text{RMS}}
\newcommand{\AARE}{\text{E}_\text{AAR}}
\newcommand{\R}{\text{R}}
\newcommand{\ps}{\text{s}^{-1}}
\DeclareMathOperator{\ANN2}{sig}
%\newcommand{\Dev}	{\mbox{\Large\ensuremath{\mathsf{s}}}}
%\DeclareMathOperator{\sech}{sech}
%\DeclareMathOperator{\ANN2}{sig}

@@

A summary of the coefficients for evaluating the high-temperature flow stress prediction capability of the AISI P20 alloy for all models presented in this work is reported in Table \ref{tab:Errors}.
\begin{table}[h!]
\centering
\caption{Accuracy coefficients for all the analyzed models.}
\begin{tabular}{lcccccc}
\hline
Coefficients & JC & MZA & HS & Arhenius & PTM & ANN\\
\hline
$\AARE(\%)$ & $14.05$ & $21.20$ & $7.75$ & $3.56$ & $4.79$ & $0.62$\\
$\RMSE(\text{MPa})$ & $12.00$ & $19.57$ & $3.80$ & $2.18$ & $4.59$ & $0.38$\\
\hline
\label{tab:Errors}
\end{tabular}
\end{table}
From this table, we can see that the ANN model has a much better predictive capacity than all the analytical models presented previously.The values of $\AARE$ and $\RMSE$ are globally $6$ times lower than the best of the analytical models, \ie the Arrhenius model, quoted as reference in the context of hot forming of alloys \cite{Liang-2022}.

@@

{Figures/inCombinaison}

@@

\begin{tabular}{lcccccc}

@@

&		&		&         &             &		   &	\\
Coefficients&JC  & MZA  &HS  & Arhenius      & PTM  &ANN \\
&				&				&         &             &	&\\

@@

$R(\%)$&$99.0590$&$99.0821$&$97.6008$&$99.8417$& $99.8760$&$99.9439$  \\
$\AARE(\%)$&$4.2686$&$3.7440$&$6.1391$&$2.3321$&$2.7590$&$1.4018$   \\
$\RMSE(\text{MPa})$&$4.1923$&$3.1856$&$82.6956$&$2.29891$&$2.4632$&$1.0228$ \\

@@

$R(\%)$&$99.2920$&$99.4879$&$93.9429$&$99.6870$& $99.7582$&$99.8047$\\
$\AARE(\%)$&$4.8988$&$3.4506$&$12.9939$&$3.2008$&$3.8845$&$1.7433$\\
$RMSE(\text{MPa})$&$5.0943$&$4.3048$&$116.517$&$3.7035$&$4.3915$&$1.2513$ \\

@@

{Figures/exCombinaison}

@@

\section*{Appendix\label{sec:Appendix}}
In order to complete this paper, we report here after the computing process and the $180$ coefficients of the Artificial Neural Network ANN-3-15-7-1-sigmoid model used in Section \ref{sec:ANNmodel}.
In order to use this model we describe here after the procedure to compute the flow stress $\sigma^y$ from the input variables $\varepsilon^p$, $\mdot\varepsilon$ and $T$.
This process can be decomposed in $3$ phases:
\begin{itemize}
\item We first have to normalize the input values of the ANN $x_i$ within the range $[0,1]$ to avoid ill-conditioned system as presented by many other authors in the literature \cite{Lin-2008-ANN, Lu-2011-ANN}.
Therefore, the three components of the input are obtained from the plastic strain $\varepsilon^p$, the plastic strain rate $\mdot{\varepsilon}^p$ and the temperature $T$ using the following expressions:
\begin{equation}
\begin{cases}
x_1 = \frac{\varepsilon^p - [\varepsilon^p]_{min}}{[\varepsilon^p]_{max} - [\varepsilon^p]_{min}}\\
x_2 = \frac{\ln(\mdot{\varepsilon}/\mdot{\varepsilon_0})-[\ln(\mdot{\varepsilon}/\mdot{\varepsilon_0})]_{min}}{[\ln(\mdot{\varepsilon}/\mdot{\varepsilon_0})]_{max}-[\ln(\mdot{\varepsilon}/\mdot{\varepsilon_0})]_{min}}\label{eq:CR1}\\
x_3 = \frac{T-[T]_{min}}{[T]_{max}-[T]_{min}}
\end{cases}
\end{equation}
where $[~]_{min}$ and $[~]_{max}$  are the boundaries of the range of the corresponding field: $\varepsilon^p\!\in\!\left[0.0,0.7\right]$, $\mdot{\varepsilon}\!\in\!\left[0.001,5.0\right]$, $T\!\in\!\left[1050,1250\right]$ and $\sigma\!\in\!\left[1.311,153.739\right]$.
The reference strain rate is $\mdot{\varepsilon_0} = 0.001$.
\item Then we compute the output $s$ of the ANN from the input vector $\overrightarrow{x}$ using the three following equations.
\begin{equation}
\overrightarrow{y}_1 = \left[1 + \exp{\left(- \w_1 \dotp \overrightarrow{x}- \overrightarrow{b}_1\right)}\right]^{-1}
\end{equation}
\begin{equation}
\overrightarrow{y}_2 = \left[1 + \exp{\left(- \w_2 \dotp \overrightarrow{y}_1- \overrightarrow{b}_2\right)}\right]^{-1}
\end{equation}
\begin{equation}
s = \overrightarrow{w}^T \dotp \overrightarrow{y}_2 + b
\end{equation}
\item Finally, the flow stress $\sigma^y$ can be obtained from the output $s$ of the ANN using the following equation:
\begin{equation}
\sigma^y =  \left([\sigma]_{max}-[\sigma]_{min}\right)s + [\sigma]_{min} \label{eq:CR2}
\end{equation}
\end{itemize}

Conforming to the computing process proposed by Equations (\ref{eq:CR1}-\ref{eq:CR2}), we report hereafter the $180$ coefficients of the ANN-3-15-7-1-sigmoid model used in Section \ref{sec:ANNmodel}.
The weight matrix for the first hidden layer $\w_1$ is a $15\times3$ matrix:
\begin{equation*}
\w_1 = \left[
\begin{array}{rrr}
2.2206 & -3.7555 & -6.7246\\ 
-4.8598 & 5.7431 & -5.8538\\ 
2.3099 & 3.3325 & -5.1795\\ 
2.0475 & 0.8006 & -1.4259\\ 
8.8358 & -6.0362 & 0.8226\\ 
-1.2613 & -0.9274 & -2.3725\\ 
-0.3561 & 6.5032 & -10.7573\\ 
-11.7226 & -2.0455 & 1.2248\\ 
3.1066 & 26.5580 & 18.6540\\ 
-0.5150 & -5.6922 & 1.0104\\ 
-6.4755 & 8.4888 & -2.4459\\ 
-1.8791 & -0.5380 & 2.2295\\ 
-6.0206 & 1.2776 & 0.2169\\ 
0.2619 & -4.7974 & -1.1282\\ 
-27.1456 & -0.5327 & -0.5303\\ 
\end{array}\right]
\end{equation*}
The biases of the first hidden layer $\overrightarrow{b_1}$ is a $15$ components vector:
\begin{equation*}
\overrightarrow{b}_1 = \left[
\begin{array}{r}
5.9164\\ 
-1.9074\\ 
-2.6837\\ 
-1.0954\\ 
-0.9509\\ 
3.1682\\ 
-4.3964\\ 
0.6343\\ 
-5.0676\\ 
2.0228\\ 
1.1267\\ 
-0.9671\\ 
-0.6263\\ 
2.3110\\ 
-0.3037\\ 
\end{array}\right]
\end{equation*}
The weight matrix for the second hidden layer $\w_2$ is a $7\times15$ matrix:
\begin{equation*}
\w_2^T = \left[
\begin{array}{rrrrrrr}
-0.5783 & 1.2724 & 0.5747 & 0.6449 & -4.2203 & -0.2380 & 0.2591\\ 
-0.4852 & 5.2807 & 0.8888 & -8.2324 & -2.0075 & -0.6474 & -1.0787\\ 
4.4499 & -0.0137 & 0.1657 & 0.3198 & 4.9765 & -1.2503 & 0.8219\\ 
1.7571 & 0.7730 & 0.0208 & -1.3316 & -0.8945 & -0.7284 & -0.1831\\ 
-1.0866 & 0.1330 & -0.8615 & -0.1283 & 0.2218 & -0.1772 & -2.7458\\ 
-0.3925 & 1.3994 & 0.0630 & -1.8397 & -1.1047 & -1.9839 & 0.5767\\ 
-0.2121 & 0.9977 & 1.2028 & -9.6525 & 0.5520 & 0.1062 & -0.0409\\ 
-1.1518 & -1.6402 & -4.1501 & -1.0759 & 0.4749 & -2.8350 & 0.9225\\ 
-1.1453 & -0.2173 & -0.1382 & 0.8264 & -0.5125 & 0.1882 & -0.8654\\ 
-4.3301 & -0.3711 & -7.4305 & 3.5926 & -9.6217 & -1.2375 & 1.6171\\ 
2.3907 & -1.0085 & -0.8828 & -1.1891 & 0.9947 & 1.1178 & -1.0953\\ 
-1.5955 & 1.6313 & 0.4916 & 0.1906 & -1.9216 & -1.4140 & 1.3827\\ 
-1.6985 & 1.4277 & -3.4462 & -8.3777 & -1.2132 & -1.2158 & 2.8512\\ 
-1.9954 & -2.0159 & -8.3455 & -0.5205 & 0.2942 & -1.3337 & 0.2026\\ 
4.3040 & -0.7164 & -1.0859 & 3.4294 & -23.8003 & 12.5859 & 7.3721\\ 
\end{array}\right]
\end{equation*}
The biases of the second hidden layer $\overrightarrow{b_2}$ is a $7$ components vector:
\begin{equation*}
\overrightarrow{b}_2 = \left[
\begin{array}{r}
0.7534\\ 
0.9473\\ 
0.6055\\ 
-0.7793\\ 
-1.0305\\ 
-1.5779\\ 
-0.1471\\ 
\end{array}\right]
\end{equation*}
The weight vector for the output layer $\overrightarrow{w}_3$ is a $7$ components vector:
\begin{equation*}
\overrightarrow{w}_3 = \left[
\begin{array}{r}
0.1920\\ 
0.3406\\ 
0.3839\\ 
-0.2880\\ 
1.2047\\ 
-1.4126\\ 
-0.2215\\ 
\end{array}\right]
\end{equation*}
The bias of the output layer $b_3$ is a scalar:
\begin{equation*}
b_3 = 0.1178 
\end{equation*}

>>>>

\newcommand{\w}{\mbox{\large\ensuremath{\mathsf{w}}}}
\newcommand{\dotp}{\boldsymbol{\cdot}}
%\newcommand{\ccirc}{\kern0.5ex\vcenter{\hbox{$\scriptstyle\circ$}}\kern0.5ex}
%\newcommand{\e}[1]{\exp\left({#1}\right)}
\newcommand{\e}[1]{{\rm e}^{#1}}
\newcommand{\lay}[1]{^{(#1)}}
\newcommand{\mdot}[1]{\accentset{\mbox{\bfseries .}}{#1}}
\newcommand*{\ie}{\emph{i.e.}\@\xspace}
\newcommand*{\versus}{\emph{vs.}\@\xspace}
\newcommand*{\eal}{et \emph{al.}\@\xspace}
\newcommand*{\eg}{e.g.,\@\xspace}
\newcommand{\RMSE}{\text{E}_\text{RMS}}
\newcommand{\AARE}{\text{E}_\text{AAR}}
\newcommand{\R}{\text{R}}
\newcommand{\ps}{\text{s}^{-1}}
%\DeclareMathOperator{\ANN2}{sig}
\newcommand{\mr}[2]{\multirow{#1}{*}{#2}}
%\newcommand{\Dev}	{\mbox{\Large\ensuremath{\mathsf{s}}}}
%\DeclareMathOperator{\sech}{sech}
%\DeclareMathOperator{\ANN2}{sig}

@@

A summary of the coefficients for evaluating the high-temperature flow stress prediction capability of the AISI P20 alloy for all models presented in this work is reported in Table \ref{tab:Errors}.
\begin{table}[h!]
\centering
\caption{Accuracy coefficients for all the analyzed models.}
\begin{tabular}{lcccccc}
\hline
Coefficients & JC & MZA & HS & AR & PTM & ANN\\
\hline
$\AARE(\%)$ & $14.05$ & $21.20$ & $7.75$ & $3.56$ & $4.79$ & $0.62$\\
$\RMSE(\text{MPa})$ & $12.00$ & $19.57$ & $3.80$ & $2.18$ & $4.59$ & $0.38$\\
\hline
\label{tab:Errors}
\end{tabular}
\end{table}
From this table, we can see that the ANN model has a much better predictive capacity than all the analytical models presented previously.The values of $\AARE$ and $\RMSE$ are globally $6$ times lower than the best of the analytical models, \ie the Arrhenius model, quoted as reference in the context of hot forming of alloys \cite{Liang-2022}.

@@

{Figures/CompInt}

@@

\begin{tabular}{llcccccc}

@@

$\mdot\varepsilon$           & Coefficients        & JC      & MZA     & HS     & AR     & PTM    & ANN\\

@@

\mr{2}{all}                  & $\AARE(\%)$         & $13.97$ & $21.49$ & $7.76$ & $3.54$ & $4.78$ & $0.78$\\
                             & $\RMSE(\text{MPa})$ & $11.86$ & $19.86$ & $3.84$ & $2.17$ & $4.45$ & $0.71$\\\hline
\mr{2}{$\mdot\varepsilon=1$} & $\AARE(\%)$         & $14.87$ & $27.79$ & $3.74$ & $1.46$ & $3.16$ & $1.40$\\
                             & $\RMSE(\text{MPa})$ & $12.02$ & $23.67$ & $3.10$ & $1.46$ & $2.65$ & $1.52$\\

@@

\mr{2}{all}                  & $\AARE(\%)$         & $14.27$ & $19.96$ & $7.37$ & $3.95$ & $5.53$  & $1.15$\\
                             & $\RMSE(\text{MPa})$ & $12.36$ & $18.24$ & $4.43$ & $2.68$ & $6.23$  & $2.40$\\\hline
\mr{2}{$\mdot\varepsilon=5$} & $\AARE(\%)$         & $20.16$ & $24.27$ & $7.37$ & $3.53$ & $11.46$ & $3.87$\\
                             & $\RMSE(\text{MPa})$ & $20.73$ & $25.95$ & $7.83$ & $4.02$ & $12.91$ & $5.84$\\

@@

{Figures/CompExt}

@@

\section*{Appendix\label{sec:Appendix}}
In order to complete this paper, we report here after the computing process and the $180$ coefficients of the Artificial Neural Network ANN-3-15-7-1-sigmoid model used in Section \ref{sec:ANNmodel}.
In order to use this model we describe here after the procedure to compute the flow stress $\sigma^y$ from the input variables $\varepsilon^p$, $\mdot\varepsilon$ and $T$.
This process can be decomposed in $3$ phases:
\begin{itemize}
\item We first have to normalize the input values of the ANN $x_i$ within the range $[0,1]$ to avoid ill-conditioned system as presented by many other authors in the literature \cite{Lin-2008-ANN, Lu-2011-ANN}.
Therefore, the three components of the input vector $\overrightarrow{x}$ are obtained from the plastic strain $\varepsilon^p$, the plastic strain rate $\mdot{\varepsilon}^p$ and the temperature $T$ using the following expressions:
\begin{equation}
\overrightarrow{x}=
\begin{cases}
x_1 = \frac{\varepsilon^p - [\varepsilon^p]_{min}}{[\varepsilon^p]_{max} - [\varepsilon^p]_{min}}\\
x_2 = \frac{\ln(\mdot{\varepsilon}/\mdot{\varepsilon_0})-[\ln(\mdot{\varepsilon}/\mdot{\varepsilon_0})]_{min}}{[\ln(\mdot{\varepsilon}/\mdot{\varepsilon_0})]_{max}-[\ln(\mdot{\varepsilon}/\mdot{\varepsilon_0})]_{min}}\label{eq:CR1}\\
x_3 = \frac{T-[T]_{min}}{[T]_{max}-[T]_{min}}
\end{cases}
\end{equation}
where $[~]_{min}$ and $[~]_{max}$  are the boundaries of the range of the corresponding field: $\varepsilon^p\!\in\!\left[0.0,0.7\right]$, $\mdot{\varepsilon}\!\in\!\left[0.001,5.0\right]$, $T\!\in\!\left[1050,1250\right]$ and $\sigma\!\in\!\left[1.311,153.739\right]$.
The reference strain rate is $\mdot{\varepsilon_0} = 0.001$.
\item Then we compute the output $s$ of the ANN from the input vector $\overrightarrow{x}$ using the three following equations.
\begin{equation}
\overrightarrow{y}_1 = \left[1 + \exp{\left(- \w_1 \dotp \overrightarrow{x}- \overrightarrow{b}_1\right)}\right]^{-1}
\end{equation}
\begin{equation}
\overrightarrow{y}_2 = \left[1 + \exp{\left(- \w_2 \dotp \overrightarrow{y}_1- \overrightarrow{b}_2\right)}\right]^{-1}
\end{equation}
\begin{equation}
s = \overrightarrow{w}^T \dotp \overrightarrow{y}_2 + b
\end{equation}
\item Finally, the flow stress $\sigma^y$ can be obtained from the output $s$ of the ANN using the following equation:
\begin{equation}
\sigma^y =  \left([\sigma]_{max}-[\sigma]_{min}\right)s + [\sigma]_{min} \label{eq:CR2}
\end{equation}
\end{itemize}

Conforming to the computing process proposed by Equations (\ref{eq:CR1}-\ref{eq:CR2}), we report hereafter the $180$ coefficients of the ANN-3-15-7-1-sigmoid model used in Section \ref{sec:ANNmodel}.
The weight matrix for the first hidden layer $\w_1$ is a $15\times3$ matrix:
\begin{equation*}
\w_1 = \left[
\begin{array}{rrr}
2.2206 & -3.7555 & -6.7246\\ 
-4.8598 & 5.7431 & -5.8538\\ 
2.3099 & 3.3325 & -5.1795\\ 
2.0475 & 0.8006 & -1.4259\\ 
8.8358 & -6.0362 & 0.8226\\ 
-1.2613 & -0.9274 & -2.3725\\ 
-0.3561 & 6.5032 & -10.7573\\ 
-11.7226 & -2.0455 & 1.2248\\ 
3.1066 & 26.5580 & 18.6540\\ 
-0.5150 & -5.6922 & 1.0104\\ 
-6.4755 & 8.4888 & -2.4459\\ 
-1.8791 & -0.5380 & 2.2295\\ 
-6.0206 & 1.2776 & 0.2169\\ 
0.2619 & -4.7974 & -1.1282\\ 
-27.1456 & -0.5327 & -0.5303\\ 
\end{array}\right]
\end{equation*}
The biases of the first hidden layer $\overrightarrow{b_1}$ is a $15$ components vector:
\begin{equation*}
\overrightarrow{b}_1 = \left[
\begin{array}{r}
5.9164\\ 
-1.9074\\ 
-2.6837\\ 
-1.0954\\ 
-0.9509\\ 
3.1682\\ 
-4.3964\\ 
0.6343\\ 
-5.0676\\ 
2.0228\\ 
1.1267\\ 
-0.9671\\ 
-0.6263\\ 
2.3110\\ 
-0.3037\\ 
\end{array}\right]
\end{equation*}
The weight matrix for the second hidden layer $\w_2$ is a $7\times15$ matrix:
\begin{equation*}
\w_2^T = \left[
\begin{array}{rrrrrrr}
-0.5783 & 1.2724 & 0.5747 & 0.6449 & -4.2203 & -0.2380 & 0.2591\\ 
-0.4852 & 5.2807 & 0.8888 & -8.2324 & -2.0075 & -0.6474 & -1.0787\\ 
4.4499 & -0.0137 & 0.1657 & 0.3198 & 4.9765 & -1.2503 & 0.8219\\ 
1.7571 & 0.7730 & 0.0208 & -1.3316 & -0.8945 & -0.7284 & -0.1831\\ 
-1.0866 & 0.1330 & -0.8615 & -0.1283 & 0.2218 & -0.1772 & -2.7458\\ 
-0.3925 & 1.3994 & 0.0630 & -1.8397 & -1.1047 & -1.9839 & 0.5767\\ 
-0.2121 & 0.9977 & 1.2028 & -9.6525 & 0.5520 & 0.1062 & -0.0409\\ 
-1.1518 & -1.6402 & -4.1501 & -1.0759 & 0.4749 & -2.8350 & 0.9225\\ 
-1.1453 & -0.2173 & -0.1382 & 0.8264 & -0.5125 & 0.1882 & -0.8654\\ 
-4.3301 & -0.3711 & -7.4305 & 3.5926 & -9.6217 & -1.2375 & 1.6171\\ 
2.3907 & -1.0085 & -0.8828 & -1.1891 & 0.9947 & 1.1178 & -1.0953\\ 
-1.5955 & 1.6313 & 0.4916 & 0.1906 & -1.9216 & -1.4140 & 1.3827\\ 
-1.6985 & 1.4277 & -3.4462 & -8.3777 & -1.2132 & -1.2158 & 2.8512\\ 
-1.9954 & -2.0159 & -8.3455 & -0.5205 & 0.2942 & -1.3337 & 0.2026\\ 
4.3040 & -0.7164 & -1.0859 & 3.4294 & -23.8003 & 12.5859 & 7.3721\\ 
\end{array}\right]
\end{equation*}
The biases of the second hidden layer $\overrightarrow{b_2}$ is a $7$ components vector:
\begin{equation*}
\overrightarrow{b}_2 = \left[
\begin{array}{r}
0.7534\\ 
0.9473\\ 
0.6055\\ 
-0.7793\\ 
-1.0305\\ 
-1.5779\\ 
-0.1471\\ 
\end{array}\right]
\end{equation*}
The weight vector for the output layer $\overrightarrow{w}$ is a $7$ components vector:
\begin{equation*}
\overrightarrow{w} = \left[
\begin{array}{r}
0.1920\\ 
0.3406\\ 
0.3839\\ 
-0.2880\\ 
1.2047\\ 
-1.4126\\ 
-0.2215\\ 
\end{array}\right]
\end{equation*}
The bias of the output layer $b$ is a scalar:
\begin{equation*}
b = 0.1178 
\end{equation*}
